---
title: "Explore 2"
author: "Blain Morin"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10)

library(tidyverse)
library(haven)
library(MatchIt)
library(plotly)
wide.df =  read_dta("Combined_Sandy_Harvey-Oct-08-2019-wide.dta")
long.df =  read_dta("Combined_Sandy_Harvey-Oct-08-2019-long.dta")

long.df = long.df %>%
  filter(all_outliers == 0)

wide.df = wide.df %>%
  filter(all_outliers == 0) %>%
  mutate(industry_codes = as.factor(industry_codes))


df = long.df %>%
  drop_na(cost_sav) %>%
  drop_na(av_loss) %>%
  mutate(Name_of_Tactic = as.factor(Name_of_Tactic)) %>%
  mutate(recovery_shape = as.factor(recovery_path_Combined)) %>%
  mutate(hurricane = as.factor(hurricane)) %>%
  mutate(bi_dummy_employeescantwork = as.factor(ifelse(bi_employees_unable_to_work > 0, 1, 0))) %>%
  mutate(bi_dummy_employeesmoved = as.factor(ifelse(bi_employees_moved_away > 0 , 1, 0))) %>%
  mutate(bi_dummy_comms = as.factor(ifelse(bi_communications > 0, 1, 0))) %>%
  mutate(bi_dummy_supplychain = as.factor(ifelse(bi_supply_chain > 0, 1, 0))) %>%
  mutate(bi_dummy_powerout = as.factor(ifelse(bi_power_outages > 0 , 1, 0))) %>%
  mutate(bi_dummy_natgas = as.factor(ifelse(bi_natural_gas_outages > 0, 1, 0))) %>%
  mutate(bi_dummy_water = as.factor(ifelse(bi_water_outages > 0, 1, 0))) %>%
  mutate(bi_dummy_transport = as.factor(ifelse(bi_transportation > 0, 1, 0))) %>%
  mutate(log_prop_damage = log(property_damage_estimate + 1)) %>%
  mutate(log_prior_emp = log(employees_prior_hurricane)) %>%
  mutate(vrec = as.factor(vrec)) %>%
  rename(Recovered = Censor) %>%
  mutate(industry_codes = as.factor(industry_codes))

### Get Cost Incurring

df2 = df %>%
  filter(cost_sav > 0) %>%
  mutate(BCR = av_loss / cost_sav + .1) %>%
  mutate(av_loss_blain = av_loss + 1) %>%
  mutate(log_cost_sav = log(cost_sav)) %>%
  mutate(log_av_loss_blain = log(av_loss_blain)) %>%
  mutate(logBCR = log(BCR)) %>%
  filter(BCR < 50000) %>%
  filter(av_loss_blain<1000000) %>%
  filter(cost_sav<1000000)


### Get cost saving

df3 = df %>%
  filter(cost_sav < 0) %>%
  mutate(BCR = av_loss / abs(cost_sav) + .1) %>%
  mutate(av_loss_blain = av_loss + 1) %>%
  mutate(log_cost_sav = log(abs(cost_sav))) %>%
  mutate(log_av_loss_blain = log(av_loss_blain)) %>%
  mutate(logBCR = log(BCR)) %>%
  filter(BCR < 50000)%>%
  filter(av_loss_blain<1000000) %>%
  filter(cost_sav<1000000)

```

# Loss share

* We develop the loss share (LS) metric to answer the question: Are losses from some disruption types less avoidable?

The loss share of disruption $d$ for firm $i$ is:

$$ LS_{d,i} = \frac{BI_{d,i}\% * LostSR_i}{NoDisasterSR_i} - \frac{BI_{d,i}\% * MaxPotentialLoss_i}{No DisasterSR_i}  $$
Which is equal to:

$$ LS_{d,i} = BI_{d,i} * RM_i  $$

* ***Interpretation***: A higher LS means that the firm experiencing the disruption was able to avoid more of its maximum potential losses. 

```{r}

df4 = df %>%
  group_by(vrec) %>%
  mutate(Total_AvLoss = sum(av_loss)) %>%
  ungroup()

df5 = df4 %>%
  mutate(RM_Total = Total_AvLoss / max_potential_losses) %>%
  mutate(LS_emp_cant_work = bi_employees_unable_to_work / 100 * RM_Total) %>%
  mutate(LS_moved_away = bi_employees_moved_away / 100 * RM_Total) %>%
  mutate(LS_communications = bi_communications  / 100 * RM_Total) %>%
  mutate(LS_Supply_Chain = bi_supply_chain / 100 * RM_Total) %>%
  mutate(LS_Transportation = bi_transportation / 100 * RM_Total) %>%
  mutate(LS_Power = bi_power_outages / 100 * RM_Total) %>%
  mutate(LS_Gas = bi_natural_gas_outages / 100 * RM_Total) %>%
  mutate(LS_Water = bi_water_outages / 100 * RM_Total)

df6 = df5 %>%
  group_by(vrec) %>%
  slice(1) %>%
  ungroup() %>%
  select(vrec, industry_codes, starts_with("LS"))


df6 = df6 %>%
  gather(LS_Type, LS, LS_emp_cant_work:LS_Water)

df6 %>%
  ggplot(aes(x = LS_Type, y = LS)) +
  geom_boxplot() +
  ylab("Loss Share") +
  xlab("Disruption Type") +
  ggtitle("Observed Loss Shares") +
  theme_classic()



```



```{r}

df6 %>%
  filter(LS_Type %in% c('LS_Power', 'LS_Gas', 'LS_Water')) %>%
  ggplot(aes(x = LS_Type, y = LS)) +
  geom_boxplot(aes(fill = industry_codes)) +
  ylab("Loss Share") +
  xlab("Disruption Type") +
  ggtitle("Observed Loss Shares From Infrastructure, By Industry") +
  theme_classic() + 
  scale_fill_brewer(palette="Paired", name = "Industry",
                      labels = c("Agriculture, Mining,\nand Construction",
                               "Transportation, Communications,\nand Utilities",
                               "Manufacturing",
                               "Wholesale, Retail, Trade",
                               "Finance, Insurance,\nand Real Estate",
                               "Service Sector")) +
  theme(legend.position="top")

```

# Loss Triangle


* *Possible paper direction* compare analyses between loss triangle metrics and max potential loss metrics. 
* Question: I'm not sure what to use for:

$$ LossTriangle = \frac{BI\%*LostSR_I}{SRNoDisaster_i}  $$

Is it: If Sandy, use month after and if Harvey use SR lost during recovery or SR lost year after?

## Figures

These tells us what proportion of the loss triangle is due to each of the disruptions. 

```{r}

df7 = df %>%
  mutate(lost_nod = ifelse(hurricane == 0 ,month_lost_sales__percent, annual_lost_sales__percent)) %>%
  mutate(lost_nod = lost_nod / 100) %>%
  mutate(LT_emp_cant_work = bi_employees_unable_to_work / 100 * lost_nod) %>%
  mutate(LT_moved_away = bi_employees_moved_away / 100 * lost_nod) %>%
  mutate(LT_communications = bi_communications  / 100 * lost_nod) %>%
  mutate(LT_Supply_Chain = bi_supply_chain / 100 * lost_nod) %>%
  mutate(LT_Transportation = bi_transportation / 100 * lost_nod) %>%
  mutate(LT_Power = bi_power_outages / 100 * lost_nod) %>%
  mutate(LT_Gas = bi_natural_gas_outages / 100 * lost_nod) %>%
  mutate(LT_Water = bi_water_outages / 100 * lost_nod)


```


```{r}


df8 = df7 %>%
  group_by(vrec) %>%
  slice(1) %>%
  ungroup() %>%
  select(vrec, industry_codes, starts_with("LT"))

df8 = df8 %>%
  gather(LT_Type, LT, LT_emp_cant_work:LT_Water)

df8 %>%
  ggplot(aes(x = LT_Type, y = LT)) +
  geom_boxplot() +
  ylab("Loss Triangle") +
  xlab("Disruption Type") +
  ggtitle("Observed Loss Triangles") +
  theme_classic()

```

```{r}

df8 %>%
  filter(LT_Type %in% c('LT_Power', 'LT_Gas', 'LT_Water')) %>%
  ggplot(aes(x = LT_Type, y = LT)) +
  geom_boxplot(aes(fill = industry_codes)) +
  ylab("Loss Triangle") +
  xlab("Disruption Type") +
  ggtitle("Observed Loss Triangles From Infrastructure, By Industry") +
  theme_classic() + 
  scale_fill_brewer(palette="Paired", name = "Industry",
                      labels = c("Agriculture, Mining,\nand Construction",
                               "Transportation, Communications,\nand Utilities",
                               "Manufacturing",
                               "Wholesale, Retail, Trade",
                               "Finance, Insurance,\nand Real Estate",
                               "Service Sector")) +
  theme(legend.position="top")

```