---
title: "Exploratory Analysis"
author: "Blain Morin"
date: "8/5/2021"
output: html_document
---



```{r setup, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10)

library(tidyverse)
library(haven)
library(plotly)
wide.df =  read_dta("Combined_Sandy_Harvey-Oct-08-2019-wide.dta")
long.df =  read_dta("Combined_Sandy_Harvey-Oct-08-2019-long.dta")

long.df = long.df %>%
  filter(all_outliers == 0)

wide.df = wide.df %>%
  filter(all_outliers == 0) %>%
  mutate(industry_codes = as.factor(industry_codes))


df = long.df %>%
  drop_na(cost_sav) %>%
  drop_na(av_loss) %>%
  mutate(Name_of_Tactic = as.factor(Name_of_Tactic)) %>%
  mutate(recovery_shape = as.factor(recovery_path_Combined)) %>%
  mutate(hurricane = as.factor(hurricane)) %>%
  mutate(bi_dummy_employeescantwork = as.factor(ifelse(bi_employees_unable_to_work > 0, 1, 0))) %>%
  mutate(bi_dummy_employeesmoved = as.factor(ifelse(bi_employees_moved_away > 0 , 1, 0))) %>%
  mutate(bi_dummy_comms = as.factor(ifelse(bi_communications > 0, 1, 0))) %>%
  mutate(bi_dummy_supplychain = as.factor(ifelse(bi_supply_chain > 0, 1, 0))) %>%
  mutate(bi_dummy_powerout = as.factor(ifelse(bi_power_outages > 0 , 1, 0))) %>%
  mutate(bi_dummy_natgas = as.factor(ifelse(bi_natural_gas_outages > 0, 1, 0))) %>%
  mutate(bi_dummy_water = as.factor(ifelse(bi_water_outages > 0, 1, 0))) %>%
  mutate(bi_dummy_transport = as.factor(ifelse(bi_transportation > 0, 1, 0))) %>%
  mutate(log_prop_damage = log(property_damage_estimate + 1)) %>%
  mutate(log_prior_emp = log(employees_prior_hurricane)) %>%
  mutate(vrec = as.factor(vrec)) %>%
  rename(Recovered = Censor) %>%
  mutate(industry_codes = as.factor(industry_codes))

df2 = df %>%
  filter(cost_sav > 0) %>%
  mutate(BCR = av_loss / cost_sav + .1) %>%
  mutate(av_loss_blain = av_loss + 1) %>%
  mutate(log_cost_sav = log(cost_sav)) %>%
  mutate(log_av_loss_blain = log(av_loss_blain)) %>%
  mutate(logBCR = log(BCR)) %>%
  filter(BCR < 50000) %>%
  filter(av_loss_blain<1000000) %>%
  filter(cost_sav<1000000)


### Get cost saving

df3 = df %>%
  filter(cost_sav < 0) %>%
  mutate(BCR = av_loss / abs(cost_sav) + .1) %>%
  mutate(av_loss_blain = av_loss + 1) %>%
  mutate(log_cost_sav = log(abs(cost_sav))) %>%
  mutate(log_av_loss_blain = log(av_loss_blain)) %>%
  mutate(logBCR = log(BCR)) %>%
  filter(BCR < 50000)%>%
  filter(av_loss_blain<1000000) %>%
  filter(cost_sav<1000000)

```

# Average Disruptions by Sector

```{r}

dff = wide.df %>%
  group_by(industry_codes) %>%
  summarise("Supply Chain Disruptions" = mean(bi_supply_chain, na.rm = T),
            "Employees Unable to \nTravel to Work     " = mean(bi_employees_unable_to_work, na.rm = T),
            "Transportation Disruption" = mean(bi_transportation, na.rm = T),
            "Communication Problems" = mean(bi_communications, na.rm = T),
            "Power Outages" = mean(bi_power_outages, na.rm = T),
            "Water Outages" = mean(bi_water_outages, na.rm = T),
            "Gas Outages" = mean(bi_natural_gas_outages, na.rm = T),
            "Employees Moved Away" = mean(bi_employees_moved_away, na.rm = T)
            )

dff = dff %>%
  gather(key = BI, value = Average, "Supply Chain Disruptions":"Employees Moved Away", factor_key = T)

p = dff %>%
  mutate(industry_codes = as.factor(industry_codes)) %>%
  ggplot() +
  geom_bar(aes(x = BI, y = Average, fill = industry_codes), stat = "identity", position = position_dodge()) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, vjust = 1)) +
  scale_fill_brewer(palette="Paired", name = "Industry",
                      labels = c("Agriculture, Mining,\nand Construction",
                               "Transportation, Communications,\nand Utilities",
                               "Manufacturing",
                               "Wholesale, Retail, Trade",
                               "Finance, Insurance,\nand Real Estate",
                               "Service Sector")) +
  xlab("") +
  ylab("Average Impact (%)\non Business Interruption") 

p


```



# Observed RM Distributions

```{r}

rms = df %>%
  mutate(rm = av_loss/max_potential_losses) %>%
  group_by(vrec) %>%
  summarise(RM_Total = sum(rm), SPENT_total = sum(cost_sav), AVOIDED_total = sum(av_loss))

firm_characteristics = df %>%
  group_by(vrec) %>%
  slice_head() 

firm_characteristics = firm_characteristics[, c(1, 12, 20, 25, 27:36, 67)]

rm.df = rms %>%
  left_join(firm_characteristics) %>%
  mutate(Hurricane = ifelse(hurricane == "0", "Sandy", "Harvey"))

```

```{r, fig.width=10}

mu = rm.df %>%
  group_by(Hurricane) %>%
  summarise(Mean = mean(RM_Total))

a = rm.df %>%
  ggplot(aes(x = RM_Total)) +
  geom_density(aes(color = Hurricane), size = 1.3) +
  geom_vline(data=mu, aes(xintercept=Mean, color=Hurricane),
             linetype="dashed", size = 1.2, alpha = .8) +
  geom_density(size = 1.5) +
  theme_bw() +
  xlab("Overall Firm RM") +
  ggtitle("By Hurricane")

mu = rm.df %>%
  group_by(industry_codes) %>%
  summarise(Mean = mean(RM_Total))

b = rm.df %>%
  ggplot(aes(x = RM_Total)) +
  geom_density(aes(color = industry_codes), size = 1.3) +
  geom_vline(data = mu, aes(xintercept=Mean, color=industry_codes),
             linetype="dashed", size = 1.2, alpha = .8) +
  geom_density(size = 1.5) +
  theme_bw() +
  xlab("Overall Firm RM") +
  ggtitle("By Industry") +
  scale_color_brewer(palette="Paired", name = "Industry",
                      labels = c("Agriculture, Mining,\nand Construction",
                               "Transportation, Communications,\nand Utilities",
                               "Manufacturing",
                               "Wholesale, Retail, Trade",
                               "Finance, Insurance,\nand Real Estate",
                               "Service Sector"))

a

b

```

# RMs versus Infrastructure

```{r}

rm.df %>%
  ggplot(aes(x = bi_employees_unable_to_work, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw() +
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_employees_moved_away, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_communications, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_supply_chain, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_power_outages, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_natural_gas_outages, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_water_outages, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)

rm.df %>%
  ggplot(aes(x = bi_transportation, y = RM_Total)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_bw()+
  facet_wrap(~industry_codes)


```

# RMs versus Property Damage

```{r}

rm.df %>% 
  mutate(log.prop = log(property_damage_estimate + 1)) %>%
  ggplot(aes(y = RM_Total, x = log.prop)) +
  geom_point() +
  geom_smooth(method = "lm",formula = y ~ poly(x, 2), se = FALSE, color = "red") +
  ylab("Overall Firm RM") +
  xlab("log(Property Damage)") +
  theme_bw()

```


# Diruptions by Hurricane

```{r}

dff = wide.df %>%
  group_by(hurricane) %>%
  summarise("Supply Chain Disruptions" = mean(bi_supply_chain, na.rm = T),
            "Employees Unable to \nTravel to Work     " = mean(bi_employees_unable_to_work, na.rm = T),
            "Transportation Disruption" = mean(bi_transportation, na.rm = T),
            "Communication Problems" = mean(bi_communications, na.rm = T),
            "Power Outages" = mean(bi_power_outages, na.rm = T),
            "Water Outages" = mean(bi_water_outages, na.rm = T),
            "Gas Outages" = mean(bi_natural_gas_outages, na.rm = T),
            "Employees Moved Away" = mean(bi_employees_moved_away, na.rm = T)
            )

dff = dff %>%
  gather(key = BI, value = Average, "Supply Chain Disruptions":"Employees Moved Away", factor_key = T)

p = dff %>%
  mutate(Hurricane = ifelse(hurricane == 0, "Sandy", "Harvey")) %>%
  ggplot() +
  geom_bar(aes(x = BI, y = Average, fill = Hurricane), stat = "identity", position = position_dodge()) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, vjust = 1)) +
  xlab("") +
  ylab("Average Impact (%)\non Business Interruption") +
  scale_fill_brewer()

p

```


# RM versus BRC

```{r}

rm.df2 = rm.df %>%
  filter(SPENT_total != 0) %>%
  mutate(BCR_total = AVOIDED_total/SPENT_total)%>%
  filter(BCR_total > 0 & BCR_total < 10)


rm.df2 %>%
  ggplot(aes(x = RM_Total, y = BCR_total)) +
  geom_point(aes(color = industry_codes)) +
  geom_smooth(method = 'lm', formula = y ~ poly(x, 2), aes(color = industry_codes), se = FALSE) +
  geom_smooth(method = 'lm', formula = y~ poly(x, 2), color = 'black', se = FALSE) +
  theme_bw() +
  scale_color_brewer(palette="Paired", name = "Industry",
                      labels = c("Agriculture, Mining,\nand Construction",
                               "Transportation, Communications,\nand Utilities",
                               "Manufacturing",
                               "Wholesale, Retail, Trade",
                               "Finance, Insurance,\nand Real Estate",
                               "Service Sector"))

```


# Avoided Losses Versus Max Potential Losses

* Black line is the overall average line (Points above are higher than average, points below are lower).
* On average, firms can avoided more losses if there are more losses to avoid. This might be important for theory but need to mull it over. 

```{r}

wide.df %>%
  ggplot(aes(x = log(max_potential_losses), y = log(avoided_losses))) +
  geom_point(aes(color = industry_codes, size = ln_employees), alpha = .2, show.legend = FALSE) +
  geom_smooth(aes(color = industry_codes), se = FALSE, method = 'lm') +
  theme_bw() + 
  scale_color_brewer(palette="Paired", name = "Industry",
                      labels = c("Agriculture, Mining,\nand Construction",
                               "Transportation, Communications,\nand Utilities",
                               "Manufacturing",
                               "Wholesale, Retail, Trade",
                               "Finance, Insurance,\nand Real Estate",
                               "Service Sector")) +
  geom_smooth(se = FALSE, method = 'lm', color = "black", size = 2)

```

 




```{r}

df2 %>%
  filter(BCR < 21) %>%
  filter(Name_of_Tactic == 10) %>%
  ggplot(aes(y = BCR)) +
  geom_jitter(aes(x = bi_employees_unable_to_work, colour = "Employees Unable to Work"),  alpha = .5) +
  geom_smooth(aes(x = bi_employees_unable_to_work), method = "lm", se = F, color = "blue") +
  geom_jitter(aes(x = bi_communications, colour = "Communications"),  alpha = .5) +
  geom_smooth(aes(x = bi_communications), method = "lm", se = F, color = "red") +
  geom_jitter(aes(x = bi_supply_chain, colour = "Supply Chain"), alpha = .5) +
  geom_smooth(aes(x = bi_supply_chain), method = "lm", se = F, color = "green") +
  geom_jitter(aes(x = bi_employees_moved_away, colour = "Employees Moved Away"),  alpha = .5) +
  geom_smooth(aes(x = bi_employees_moved_away), method = "lm", se = F, color = "purple") +
  theme_classic() +

  ylab("Benefit Cost Ratio (Higher = More Effective)") +
  ggtitle("Cost Effectivess of Production Recapture") +
  xlab("% Business Interruption")


```
